---
title: "Comparison of Extraction Methods (proteins)"
subtitle: "all tissues (cells, fresh-frozen, FFPE, plasma, serum)"
author:
    - Thomas Naake^[European Molecular Biology Laboratory, Meyerhofstrasse 1, 69117 Heidelberg, Germany]
fig_width: 15
fig_height: 10
fontsize: 12pt
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: TRUE
    theme: united
    number_sections: true
    highlight: tango
---

```{r env, include=FALSE, echo=FALSE, cache=FALSE}
library("knitr")
library("ggplot2")
library("tidyverse")
knitr::opts_chunk$set(stop_on_error = 1L, fig.pos = "ht", dev = "png")
suppressPackageStartupMessages(library("wesanderson"))
suppressPackageStartupMessages(library("MatrixQCvis"))
suppressPackageStartupMessages(library("MatrixQCvisUtils"))
suppressPackageStartupMessages(library("PhysicoChemicalPropertiesProtein"))
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("limma"))
knitr::opts_knit$set(root.dir = "~/GitLab/20211202_extraction_method/")
suppressMessages(
    devtools::source_gist("2a1bb0133ff568cbe28d", 
        filename = "geom_flat_violin.R"))
```


In the following section check if there is a relation between physico-chemical
properties and the observed differences between the extraction methods.

Use here the protein datasets.

# Load the data set

```{r load_data}
setwd("~/GitLab/20211202_extraction_method/")

## load cells data set
cells <- maxQuant("proteinGroups_overview.xlsx", intensity = "LFQ", 
    sheet = "cells", type = "xlsx")
cells <- assay(cells) %>% 
    transformAssay(method = "log") %>%
    MatrixQCvis:::updateSE(cells, assay = .)

## load fresh-frozen data set
ff <- maxQuant("proteinGroups_overview.xlsx", intensity = "LFQ", 
    sheet = "Fresh-frozen", type = "xlsx")
ff <- assay(ff) %>% 
    transformAssay(method = "log") %>%
    MatrixQCvis:::updateSE(ff, assay = .)

## load FFPE data set
ffpe <- maxQuant("proteinGroups_overview.xlsx", intensity = "LFQ", 
    sheet = "FFPE", type = "xlsx")
ffpe <- assay(ffpe) %>% 
    transformAssay(method = "log") %>%
    MatrixQCvis:::updateSE(ffpe, assay = .)

## load Serum-plasma
plasma_serum <- maxQuant("proteinGroups_overview.xlsx", intensity = "LFQ", 
    sheet = "Serum-Plasma", type = "xlsx")
plasma_serum <- assay(plasma_serum) %>% 
    transformAssay(method = "log") %>%
    MatrixQCvis:::updateSE(plasma_serum, assay = .)
```

```{r load_annot}
column_keep <- c("Sample_IDs", "condition", "LFQ")

## load the annotation
annot <- openxlsx::read.xlsx("proteinGroups_overview.xlsx", 
    sheet = "annotation", startRow = 2)
annot <- annot[, column_keep]
annot <- annot[!annot$condition %in% "ignore", ]
annot <- annot[!is.na(annot$condition), ]

## truncate the annotation for Cells
annot_cells <- annot[grep(annot$condition, pattern = "Cells"), ]

## truncate the annotation for Fresh-frozen
annot_ff <- annot[grep(annot$Sample_IDs, pattern = "powder_tissue_AFA_|_T"), ]

## truncate the annotation for FFPE
annot_ffpe <- annot[grep(annot$Sample_IDs, pattern = "_FFPE"), ]

## truncate the annotation for Plasma and Serum
annot_plasma <- annot[grep(annot$condition, pattern = "Plasma"), ]
annot_serum <- annot[grep(annot$condition, pattern = "Serum"), ]
```

Load the protein FASTA files and cut the sequence names.

```{r biostrings, message = FALSE}
library(Biostrings)
hs <- readAAStringSet("2021_03_30_Uniprot_homo_sapiens_canonical - Copy.fasta")
mm <- readAAStringSet("20190204_uniprot-mus+musculus-filtered-reviewed_yes.fasta")

## for cells, ffpe, plasma, serum = human
names_hs <- names(hs@ranges)
names_hs <- strsplit(names_hs, split = "[|]")
names_hs <- unlist(lapply(names_hs, "[", 2))
hs <- data.frame(feature = names_hs, 
    sequence = as.character(unlist(lapply(hs, as.character))))

## for FF = mouse
names_mm <- names(mm@ranges)
names_mm <- strsplit(names_mm, split = "[|]")
names_mm <- unlist(lapply(names_mm, "[", 2))
mm <- data.frame(feature = names_mm, 
    sequence = as.character(unlist(lapply(mm, as.character))))
```

## Update the colData and rowData slots with the annotations

Write the annotations and the sequences to the `colData` and `rowData` slots.

```{r updata_cD, echo = TRUE}
## cells
annot_cells$Sample_IDs <- make.names(annot_cells$Sample_IDs)
cells <- cells[, annot_cells$Sample_IDs]
cD <- colData(cells) %>% as.data.frame()
rD <- rowData(cells) %>% as.data.frame()
rD$feature_cut <- unlist(lapply(strsplit(rD$feature, split = ";"), "[", 1))
cells@colData <- left_join(cD, annot_cells, by = c("name" = "Sample_IDs")) %>%
    DataFrame()
rowData(cells) <- left_join(rD, hs, by = c("feature_cut" = "feature")) %>% 
    DataFrame()

## fresh-frozen
annot_ff$Sample_IDs <- make.names(annot_ff$Sample_IDs)
ff <- ff[, annot_ff$Sample_IDs]
cD <- colData(ff) %>% as.data.frame()
rD <- rowData(ff) %>% as.data.frame()
rD$feature_cut <- unlist(lapply(strsplit(rD$feature, split = ";"), "[", 1))
ff@colData <- left_join(cD, annot_ff, by = c("name" = "Sample_IDs")) %>%
  DataFrame()
rowData(ff) <- left_join(rD, mm, by = c("feature_cut" = "feature")) %>% 
    DataFrame()

## FFPE
annot_ffpe$Sample_IDs <- make.names(annot_ffpe$Sample_IDs)
ffpe <- ffpe[, annot_ffpe$Sample_IDs]
cD <- colData(ffpe) %>% as.data.frame()
rD <- rowData(ffpe) %>% as.data.frame()
rD$feature_cut <- unlist(lapply(strsplit(rD$feature, split = ";"), "[", 1))
ffpe@colData <- left_join(cD, annot_ffpe, by = c("name" = "Sample_IDs")) %>%
  DataFrame()
rowData(ffpe) <- left_join(rD, hs, by = c("feature_cut" = "feature")) %>% 
    DataFrame()

## Plasma
annot_plasma$Sample_IDs <- make.names(annot_plasma$Sample_IDs)
plasma <- plasma_serum[, annot_plasma$Sample_IDs]
cD <- colData(plasma) %>% as.data.frame()
rD <- rowData(plasma) %>% as.data.frame()
rD$feature_cut <- unlist(lapply(strsplit(rD$feature, split = ";"), "[", 1))
plasma@colData <- left_join(cD, annot_plasma, by = c("name" = "Sample_IDs")) %>%
  DataFrame()
rowData(plasma) <- left_join(rD, hs, by = c("feature_cut" = "feature")) %>% 
    DataFrame()

## Serum
annot_serum$Sample_IDs <- make.names(annot_serum$Sample_IDs)
serum <- plasma_serum[, annot_serum$Sample_IDs]
cD <- colData(serum) %>% as.data.frame()
rD <- rowData(serum) %>% as.data.frame()
rD$feature_cut <- unlist(lapply(strsplit(rD$feature, split = ";"), "[", 1))
serum@colData <- left_join(cD, annot_serum, by = c("name" = "Sample_IDs")) %>%
  DataFrame()
rowData(serum) <- left_join(rD, hs, by = c("feature_cut" = "feature")) %>% 
    DataFrame()
```

# Differential expression analysis

For the differential expression analysis use in all cases the following scheme:

  - create the model matrix,
  - fit a linear model for each protein taking into account the model matrix 
    using `lmFit` and method `"ls"`,
  - create the contrasts between two levels as specified in each section,
  - compute contrasts from linear model fit; given the linear model fit to the
    data, compute estimated coefficients and standard errors for a given 
    set of contrasts  (using `contrasts.fit`)
  - run Empirical Bayes Statistics for differential expressino; 
    Given a linear model fit from `lmFit`, compute moderated t-statistics, 
    moderated F-statistic, and log-odds of differential expression by 
    empirical Bayes moderation of the standard errors towards a global value
    (using `eBayes`)

```{r go_fct, message=FALSE, warning=FALSE, echo = FALSE}
## load the packages
library("org.Hs.eg.db")
library("org.Mm.eg.db")

## define the significance level for enrichment
signProts <- function(allScore) {
   return(allScore <= 0.05)
}

## all takes the proteins into account with p < 0.05
## up takes the proteins into account with p < 0.05 AND t > 0
## down takes the proteins into account with p < 0.05 AND t < 0
enrichment <- function(tT, organism = "Hs", de = c("all", "up", "down")) {
    
    de <- match.arg(de)
    ## translate from UNIPROT into SYMBOL
    ## conversion between id's 
    if (organism == "Hs") {
        uniprots <- Rkeys(org.Hs.egUNIPROT)
        dict <- AnnotationDbi::select(org.Hs.eg.db, uniprots, "ENTREZID", "UNIPROT")
        species <- "Hs"
    }
    if (organism == "Mm") {
        uniprots <- Rkeys(org.Mm.egUNIPROT)
        dict <- AnnotationDbi::select(org.Mm.eg.db, uniprots, "ENTREZID", "UNIPROT")
        species <- "Mm"
    }
    uniprot <- dict$UNIPROT
    entrez <- dict$ENTREZID
    names(entrez) <- uniprot
    
    ## some proteins have multiple uniprot ids, only keep the first
    rownames(tT) <- unlist(
        lapply(strsplit(rownames(tT), split = "[;]"), "[", 1))
    
    ## translate into entrez
    tT_entrez <- as.character(entrez[rownames(tT)])
    tT <- tT[!is.na(tT_entrez), ]
    tT_entrez <- tT_entrez[!is.na(tT_entrez)]
    tT <- tT[!duplicated(tT_entrez), ]
    tT_entrez <- tT_entrez[!duplicated(tT_entrez)]
    rownames(tT) <- tT_entrez
    
    ## use the p-values from the raw-intensity derived values
    tT <- tT[!is.na(tT[, "adj.P.Val"]), ]
    
    if (de == "all") {
        tT_de <- tT[tT[, "adj.P.Val"] < 0.05, ]  
    }
    if (de == "up") {
        tT_de <- tT[tT[, "adj.P.Val"] < 0.05 & tT[, "t"] > 0, ]  
    }
    if (de == "down") {
        tT_de <- tT[tT[, "adj.P.Val"] < 0.05 & tT[, "t"] < 0, ]  
    }
    
    allRes <- goana(de = rownames(tT_de), 
          universe = rownames(tT), species = species)
    
    ## return the results
    allRes
}
```

```{r parameters_DE}
## set parameters for differential expression (num only for display)
num <- Inf
p_val <- 1
adj <- "BH"
```


# Cells

contrasts:
  
  - Cells_AFA - Cells_TwoPhase_AFA

## Cells: Overlap, uniqueness of features 

The plot shows the interaction of sets between different variables depending on
their presence. The dots in the UpSet plot specify if the criteria for presence
are fulfilled: Presence is defined by a feature being measured in at least one
sample of a set.

```{r overlap_cells, echo = FALSE}
MatrixQCvis::upsetCategory(cells, category = "condition", measured = TRUE)

pdf("upset_protein_cells_all.pdf")
MatrixQCvis::upsetCategory(cells, category = "condition", measured = TRUE)
dev.off()

feat_cells <- extractComb(cells, combination = c("Cells_AFA", "Cells_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
```

Only continue with the shared features in the following analyses.

```{r cells_feat, echo = FALSE}
## write also the unique proteins to files
feat <- extractComb(cells, combination = c("Cells_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_cells_Cells_AFA.txt",
    quote = FALSE, row.names = FALSE)
feat <- extractComb(cells, combination = c("Cells_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_cells_Cells_TwoPhase_AFA.txt",
    quote = FALSE, row.names = FALSE)

## continue only with the overlap
feat <- extractComb(cells, combination = c("Cells_AFA", "Cells_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
cells <- cells[feat, ]
```

## Cells: CV

```{r cells_cv, echo = FALSE}
cv_a <- assay(cells)[, cells$condition == "Cells_AFA"]
cv_b <- assay(cells)[, cells$condition == "Cells_TwoPhase_AFA"]
partial_bundle(MatrixQCvis::cvFeaturePlot(list(Cells = cv_a, Cells_TwoPhase = cv_b)))

cv_a <- MatrixQCvis::cv(t(cv_a))[[1]]
cv_b <- MatrixQCvis::cv(t(cv_b))[[1]]

## check if one of the two methods has higher means of cv
print("mean: Cells_AFA")
mean(cv_a, na.rm = TRUE)
print("mean: Cells_TwoPhase_AFA")
mean(cv_b, na.rm = TRUE)
wilcox.test(cv_a, cv_b, paired = TRUE)
```

The TwoPhase method shows higher CV compared to the traditional AFA method.

## Cells: Differential expression analysis

```{r de_cells, echo = FALSE}
design <- model.matrix(~ 0 + condition, data = colData(cells))
colnames(design) <- make.names(colnames(design))
fit <- lmFit(object = assay(cells), design = design, method = "ls")

## create contrasts
contrasts <- makeContrasts(
    AFAvsTwoPhaseAFA = conditionCells_AFA - conditionCells_TwoPhase_AFA,
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

## contrast: Cells_AFA - Cells_TwoPhase_AFA
tT <- topTable(fit_eB, number = Inf, p.value = p_val, adjust.method = adj,
    coef = "AFAvsTwoPhaseAFA")
tT <- tT[feat, ]
write.table(tT, file = "de_cells_Cells_AFA_vs_Cells_Two_Phase_AFA.txt", sep = "\t", quote = FALSE)
rmarkdown::paged_table(tT[order(tT[, "adj.P.Val"]),], options = list(max.print = num))
tT <- cbind(name = rownames(tT), tT)
partial_bundle(volcanoPlot(tT))
```

## Cells: Enrichment analysis 

Run now the GO enrichment tests. Use the raw (unadjusted) values and return
the GO terms for the ontologies Biological Process (BP), Molecular Function (MF) 
and Cellular Component (CC). Perform a Fisher test with the significant features
($\alpha$ < 0.05).
```{r enrichment_cells, message=FALSE, warning=FALSE, echo = FALSE}
allRes <- enrichment(tT, organism = "Hs")
print("## BP")
rmarkdown::paged_table(topGO(allRes, ontology = "BP", number = num))
print("## MF")
rmarkdown::paged_table(topGO(allRes, ontology = "MF", number = num))
print("## CC")
rmarkdown::paged_table(topGO(allRes, ontology = "CC", number = num))
```

## Cells: Association to physico-chemical parameters

Take the t-values and plot against the GRAVY-Score and Isoelectric point

```{r cells_phyc_paramterers, echo = FALSE, eval=FALSE}
t <- tT[, "t"]

## GRAVY score
rD <- rowData(cells)
gravy <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateGravyScore(aa))
gravy <- unlist(gravy)
df <- data.frame(gravy = gravy, t = t)
ggplot(df, aes(x = gravy, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(gravy, t, method = "spearman")

## isoelectric point
iep <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateIsoelectricPoint(aa, method = "IPC_protein"))
iep <- unlist(iep)
df <- data.frame(iep = iep, t = t)
ggplot(df, aes(x = iep, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(iep, t, method = "spearman")
```


# Fresh-frozen

contrasts: 

  - powder_AFA - powder_TwoPhase_AFA (contrast 1)
  - Tissue_bulk_AFA - powder_Two_Phase_AFA (contrast 2)
  

Upset plot for all condition types.
```{r overlap_ff_all, echo = FALSE}
MatrixQCvis::upsetCategory(ff, category = "condition", measured = TRUE)

pdf("upset_protein_ff_all.pdf")
MatrixQCvis::upsetCategory(ff, category = "condition", measured = TRUE)
dev.off()

feat_ff <- extractComb(ff, 
    combination = c("powder_AFA", "powder_TwoPhase_AFA", "Tissue_bulk_AFA"), 
    measured = TRUE, category = "condition")
```

## Fresh-frozen: Overlap, uniqueness of features (contrast 1)

The plot shows the interaction of sets between different variables depending on
their presence. The dots in the UpSet plot specify if the criteria for presence
are fulfilled: Presence is defined by a feature being measured in at least one
sample of a set.

```{r overlap_ff_1, echo = FALSE}
ff_1 <- ff[, ff$condition %in% c("powder_AFA", "powder_TwoPhase_AFA")]
MatrixQCvis::upsetCategory(ff_1, category = "condition", measured = TRUE)
```

Only continue with the shared features in the following analyses.

```{r ff_1_feat, eval = TRUE, echo = FALSE}
 ## write also the unique proteins to files
feat <- extractComb(ff_1, combination = c("powder_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ff_contrast1_powder.txt",
    quote = FALSE, row.names = FALSE)
feat <- extractComb(ff_1, combination = c("powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ff_contrast1_powder_TwoPhase.txt",
    quote = FALSE, row.names = FALSE)

## continue only with the overlap
feat <- extractComb(ff_1, combination = c("powder_AFA", "powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
ff_1 <- ff_1[feat,]
```


## Fresh-frozen: CV (contrast 1)

```{r ff_1_cv, eval = TRUE, echo = FALSE}
cv_a <- assay(ff_1)[, ff_1$condition == "powder_AFA"]
cv_b <- assay(ff_1)[, ff_1$condition == "powder_TwoPhase_AFA"]
partial_bundle(MatrixQCvis::cvFeaturePlot(list(powder_AFA = cv_a, powder_TwoPhase_AFA = cv_b)))

cv_a <- MatrixQCvis::cv(t(cv_a))[[1]]
cv_b <- MatrixQCvis::cv(t(cv_b))[[1]]

## check if one of the two methods has higher means of cv
print("mean: powder_AFA")
mean(cv_a, na.rm = TRUE)
print("mean: powder_TwoPhase_AFA")
mean(cv_b, na.rm = TRUE)
wilcox.test(cv_a, cv_b, paired = TRUE)
```

## Fresh-frozen: Differential expression analysis (contrast 1)

```{r de_ff_1, echo = FALSE, eval = TRUE}
design <- model.matrix(~ 0 + condition, data = colData(ff))
colnames(design) <- make.names(colnames(design))
fit <- lmFit(object = assay(ff), design = design, method = "ls")

## create contrasts
contrasts <- makeContrasts(
    powderAFAvsPowderTwoPhaseAFA = 
        conditionpowder_AFA - conditionpowder_TwoPhase_AFA,
    TissueBulkAFAvsPowderTwoPhaseAFA = 
        conditionTissue_bulk_AFA - conditionpowder_TwoPhase_AFA,
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

print("## contrast: powder_AFA - powder_TwoPhase_AFA")
tT <- topTable(fit_eB, number = Inf, p.value = p_val, adjust.method = adj,
    coef = "powderAFAvsPowderTwoPhaseAFA")
tT <- tT[feat, ]
write.table(tT, file = "de_FF_contrast1_powder_AFA_vs_powder_TwoPhase_AFA.txt", sep = "\t", quote = FALSE)
rmarkdown::paged_table(tT[order(tT[, "adj.P.Val"]),], options = list(max.print = num))
tT <- cbind(name = rownames(tT), tT)
partial_bundle(volcanoPlot(tT))
```

## Fresh-frozen: Enrichment analysis (contrast 1)

Run now the GO enrichment tests. Use the raw (unadjusted) values and return
the GO terms for the ontologies Biological Process (BP), Molecular Function (MF) 
and Cellular Component (CC). Perform a Fisher test with the significant features
($\alpha$ < 0.05).

```{r enrichment_ff_1, message=FALSE, warning=FALSE, echo = FALSE, eval = TRUE}
allRes <- enrichment(tT, organism = "Mm")
print("## BP")
rmarkdown::paged_table(topGO(allRes, ontology = "BP", number = num))
print("## MF")
rmarkdown::paged_table(topGO(allRes, ontology = "MF", number = num))
print("## CC")
rmarkdown::paged_table(topGO(allRes, ontology = "CC", number = num))
```

## Fresh-frozen: Association to physico-chemical parameters (contrast 1)

Take the t-values and plot against the GRAVY-Score and Isoelectric point

```{r ff_1_phyco, echo = FALSE, eval = FALSE}
t <- tT[, "t"]

## GRAVY score
rD <- rowData(ff)
gravy <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateGravyScore(aa))
gravy <- unlist(gravy)
df <- data.frame(gravy = gravy, t = t)
ggplot(df, aes(x = gravy, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(gravy, t, method = "spearman")

## isoelectric point
iep <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateIsoelectricPoint(aa, method = "IPC_protein"))
iep <- unlist(iep)
df <- data.frame(iep = iep, t = t)
ggplot(df, aes(x = iep, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(iep, t, method = "spearman")
```

## Fresh-frozen: Overlap, uniqueness of features (contrast 2)

The plot shows the interaction of sets between different variables depending on
their presence. The dots in the UpSet plot specify if the criteria for presence
are fulfilled: Presence is defined by a feature being measured in at least one
sample of a set.

```{r ff_2_overlap, eval = TRUE, echo = FALSE}
ff_2 <- ff[, ff$condition %in% c("Tissue_bulk_AFA", "powder_TwoPhase_AFA")]
MatrixQCvis::upsetCategory(ff_2, category = "condition", measured = TRUE)
```

Only continue with the shared features in the following analyses.

```{r ff_2_feat, eval = TRUE, echo = FALSE}
## write also the unique proteins to files
feat <- extractComb(ff_2, combination = c("Tissue_bulk_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ff_contrast2_Tissue_bulk.txt",
    quote = FALSE, row.names = FALSE)
feat <- extractComb(ff_2, combination = c("powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ff_contrast2_powder_TwoPhase.txt",
    quote = FALSE, row.names = FALSE)

feat <- extractComb(ff_2, 
    combination = c("Tissue_bulk_AFA", "powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
ff_2 <- ff_2[feat,]
```

## Fresh-frozen: CV (contrast 2)

```{r ff_2_cv, eval = TRUE, echo = FALSE}
cv_a <- assay(ff_2)[, ff_2$condition == "Tissue_bulk_AFA"]
cv_b <- assay(ff_2)[, ff_2$condition == "powder_TwoPhase_AFA"]
partial_bundle(MatrixQCvis::cvFeaturePlot(list(Tissue_bulk = cv_a, powder_TwoPhase = cv_b)))

cv_a <- MatrixQCvis::cv(t(cv_a))[[1]]
cv_b <- MatrixQCvis::cv(t(cv_b))[[1]]

## check if one of the two methods has higher means of cv
print("mean: Tissue_bulk_AFA")
mean(cv_a, na.rm = TRUE)
print("mean: powder_TwoPhase_AFA")
mean(cv_b, na.rm = TRUE)
wilcox.test(cv_a, cv_b, paired = TRUE)
```

## Fresh-frozen: Differential expression analysis (contrast 2)

```{r de_ff_2, eval = TRUE, echo = FALSE}
print("## contrast: Tissue_bulk_AFA - powder_TwoPhase_AFA")
tT <- topTable(fit_eB, number = Inf, p.value = p_val, adjust.method = adj,
    coef = "TissueBulkAFAvsPowderTwoPhaseAFA")
tT <- tT[feat, ]
write.table(tT, file = "de_FF_contrast2_Tissue_bulk_AFA_vs_powder_TwoPhase_AFA.txt", sep = "\t", quote = FALSE)
rmarkdown::paged_table(tT[order(tT[, "adj.P.Val"]),], options = list(max.print = num))
tT <- cbind(name = rownames(tT), tT)
partial_bundle(volcanoPlot(tT))
```

## Fresh-frozen: Enrichment analysis (contrast 2)

Run now the GO enrichment tests. Use the raw (unadjusted) values and return
the GO terms for the ontologies Biological Process (BP), Molecular Function (MF)
and Cellular Component (CC). Perform a Fisher test with the significant features
($\alpha$ < 0.05).
```{r enrichment_ff_2, message=FALSE, warning=FALSE, echo = FALSE, eval = TRUE}
allRes <- enrichment(tT, organism = "Mm")
print("## BP")
rmarkdown::paged_table(topGO(allRes, ontology = "BP", number = num))
print("## MF")
rmarkdown::paged_table(topGO(allRes, ontology = "MF", number = num))
print("## CC")
rmarkdown::paged_table(topGO(allRes, ontology = "CC", number = num))
```

## Fresh-frozen: Association to physico-chemical parameters (contrast 2)

Take the t-values and plot against the GRAVY-Score and Isoelectric point

```{r ff_phyco_parameters, echo = FALSE, eval = FALSE}
t <- tT[, "t"]

## GRAVY score
rD <- rowData(ff)
gravy <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateGravyScore(aa))
gravy <- unlist(gravy)
df <- data.frame(gravy = gravy, t = t)
ggplot(df, aes(x = gravy, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(gravy, t, method = "spearman")

## isoelectric point
iep <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateIsoelectricPoint(aa, method = "IPC_protein"))
iep <- unlist(iep)
df <- data.frame(iep = iep, t = t)
ggplot(df, aes(x = iep, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(iep, t, method = "spearman")
```


# FFPE

contrasts:

  - powder_AFA - powder_TwoPhase_AFA (contrast 1)
  - FFPE_AFA - powder_TwoPhase_AFA (contrast 2)
  
Upset plot for all condition types.

```{r overlap_ffpe_all, echo = FALSE}
MatrixQCvis::upsetCategory(ffpe, category = "condition", measured = TRUE)

pdf("upset_protein_ffpe_all.pdf")
MatrixQCvis::upsetCategory(ffpe, category = "condition", measured = TRUE)
dev.off()

feat_ffpe <- extractComb(ffpe, 
    combination = c("powder_AFA", "powder_TwoPhase_AFA", "FFPE_AFA"), 
    measured = TRUE, category = "condition")
```
  
## FFPE: Overlap, uniqueness of features (contrast 1)


The plot shows the interaction of sets between different variables depending on
their presence. The dots in the UpSet plot specify if the criteria for presence
are fulfilled: Presence is defined by a feature being measured in at least one
sample of a set.

```{r overlap_ffpe_1, eval = TRUE, echo = FALSE}
ffpe_1 <- ffpe[, ffpe$condition %in% c("powder_AFA", "powder_TwoPhase_AFA")]
MatrixQCvis::upsetCategory(ffpe_1, category = "condition", measured = TRUE)
```

Only continue with the shared features in the following analyses.

```{r ffpe_1_feat, echo = FALSE}
## write also the unique proteins to files
feat <- extractComb(ffpe_1, combination = c("powder_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ffpe_contrast1_powder.txt",
    quote = FALSE, row.names = FALSE)
feat <- extractComb(ffpe_1, combination = c("powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ffpe_contrast1_TwoPhase.txt",
    quote = FALSE, row.names = FALSE)

## continue only with the overlap
feat <- extractComb(ffpe_1, combination = c("powder_AFA", "powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
ffpe_1 <- ffpe_1[feat,]
```


## FFPE: CV (contrast 1)

```{r ffpe_1_cv, echo = FALSE}
cv_a <- assay(ffpe_1)[, ffpe_1$condition == "powder_AFA"]
cv_b <- assay(ffpe_1)[, ffpe_1$condition == "powder_TwoPhase_AFA"]
partial_bundle(MatrixQCvis::cvFeaturePlot(list(powder_AFA = cv_a, powder_TwoPhase_AFA = cv_b)))

cv_a <- MatrixQCvis::cv(t(cv_a))[[1]]
cv_b <- MatrixQCvis::cv(t(cv_b))[[1]]

## check if one of the two methods has higher means of cv
print("mean: powder_AFA")
mean(cv_a, na.rm = TRUE)
print("mean: powder_TwoPhase_AFA")
mean(cv_b, na.rm = TRUE)
wilcox.test(cv_a, cv_b, paired = TRUE)
```


## FFPE: Differential expression analysis (contrast 1)

```{r de_ffpe_1, echo = FALSE}
design <- model.matrix(~ 0 + condition, data = colData(ffpe))
colnames(design) <- make.names(colnames(design))
fit <- lmFit(object = assay(ffpe), design = design, method = "ls")

## create contrasts
contrasts <- makeContrasts(
    powderAFAvsPowderTwoPhaseAFA = 
        conditionpowder_AFA - conditionpowder_TwoPhase_AFA,
    ffpeAFAvsPowderTwoPhaseAFA = 
        conditionFFPE_AFA - conditionpowder_TwoPhase_AFA,
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

print("## contrast: powder_AFA - powder_TwoPhase_AFA")
tT <- topTable(fit_eB, number = Inf, p.value = p_val, adjust.method = adj,
    coef = "powderAFAvsPowderTwoPhaseAFA")
tT <- tT[feat, ]
write.table(tT, file = "de_FFPE_contrast1_powder_AFA_vs_powder_TwoPhase_AFA.txt", sep = "\t", quote = FALSE)
rmarkdown::paged_table(tT[order(tT[, "adj.P.Val"]),], options = list(max.print = num))
tT <- cbind(name = rownames(tT), tT)
partial_bundle(volcanoPlot(tT))
```

## FFPE: Enrichment analysis (contrast 1) 

Run now the GO enrichment tests. Use the raw (unadjusted) values and return
the GO terms for the ontologies Biological Process (BP), Molecular Function (MF)
and Cellular Component (CC). Perform a Fisher test with the significant features
($\alpha$ < 0.05).
```{r enrichment_ffpe_1, message=FALSE, warning=FALSE, echo = FALSE}
allRes <- enrichment(tT, organism = "Hs")
print("## BP")
rmarkdown::paged_table(topGO(allRes, ontology = "BP", 
    number = nrow(allRes[allRes[, "Ont"] == "BP", ])))
print("## MF")
rmarkdown::paged_table(topGO(allRes, ontology = "MF", 
    number = nrow(allRes[allRes[, "Ont"] == "MF", ])))
print("## CC")
rmarkdown::paged_table(topGO(allRes, ontology = "CC", 
    number = nrow(allRes[allRes[, "Ont"] == "CC", ])))
```

## FFPE: Association to physico-chemical parameters (contrast 1)

Take the t-values and plot against the GRAVY-Score and Isoelectric point

```{r ffpe_1_phyco, echo = FALSE, eval=FALSE}
t <- tT[, "t"]

## GRAVY score
rD <- rowData(ffpe)
gravy <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateGravyScore(aa))
gravy <- unlist(gravy)
df <- data.frame(gravy = gravy, t = t)
ggplot(df, aes(x = gravy, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(gravy, t, method = "spearman")

## isoelectric point
iep <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateIsoelectricPoint(aa, method = "IPC_protein"))
iep <- unlist(iep)
df <- data.frame(iep = iep, t = t)
ggplot(df, aes(x = iep, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(iep, t, method = "spearman")
```

## FFPE: Overlap, uniqueness of features (contrast 2)

The plot shows the interaction of sets between different variables depending on
their presence. The dots in the UpSet plot specify if the criteria for presence
are fulfilled: Presence is defined by a feature being measured in at least one
sample of a set.

```{r overlap_ffpe_2, echo = FALSE}
ffpe_2 <- ffpe[, ffpe$condition %in% c("FFPE_AFA", "powder_TwoPhase_AFA")]
MatrixQCvis::upsetCategory(ffpe_2, category = "condition", measured = TRUE)
```

Only continue with the shared features in the following analyses.

```{r ffpe_2_feat, echo = FALSE}
## write also the unique proteins to files
feat <- extractComb(ffpe_2, combination = c("FFPE_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ffpe_contrast2_FFPE.txt",
    quote = FALSE, row.names = FALSE)
feat <- extractComb(ffpe_2, combination = c("powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_ffpe_contrast2_powder_TwoPhase.txt",
    quote = FALSE, row.names = FALSE)

## continue only with the overlap
feat <- extractComb(ffpe_2, combination = c("FFPE_AFA", "powder_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
ffpe_2 <- ffpe_2[feat,]
```


## FFPE: CV (contrast 2)

```{r ffpe_2_cv, echo = FALSE}
cv_a <- assay(ffpe_2)[, ffpe_2$condition == "FFPE_AFA"]
cv_b <- assay(ffpe_2)[, ffpe_2$condition == "powder_TwoPhase_AFA"]
partial_bundle(MatrixQCvis::cvFeaturePlot(list(powder_AFA = cv_a, powder_TwoPhase_AFA = cv_b)))

cv_a <- MatrixQCvis::cv(t(cv_a))[[1]]
cv_b <- MatrixQCvis::cv(t(cv_b))[[1]]

## check if one of the two methods has higher means of cv
print("mean: FFPE_AFA")
mean(cv_a, na.rm = TRUE)
print("mean: powder_TwoPhase_AFA")
mean(cv_b, na.rm = TRUE)
wilcox.test(cv_a, cv_b, paired = TRUE)
```


## FFPE: Differential expression analysis (contrast 2)

```{r de_ffpe_2, echo = FALSE}
print("## contrast: FFPE_AFA - powder_TwoPhase_AFA")
tT <- topTable(fit_eB, number = Inf, p.value = p_val, adjust.method = adj,
    coef = "ffpeAFAvsPowderTwoPhaseAFA")
tT <- tT[feat, ]
write.table(tT, file = "de_FFPE_contrast2_FFPE_AFA_vs_powder_TwoPhase_AFA.txt", sep = "\t", quote = FALSE)
rmarkdown::paged_table(tT[order(tT[, "adj.P.Val"]),], options = list(max.print = num))
tT <- cbind(name = rownames(tT), tT)
partial_bundle(volcanoPlot(tT))
```

## FFPE: Enrichment analysis (contrast 2)

Run now the GO enrichment tests. Use the raw (unadjusted) values and return
the GO terms for the ontologies Biological Process (BP), Molecular Function (MF)
and Cellular Component (CC). Perform a Fisher test with the significant features
($\alpha$ < 0.05).
```{r enrichment_ffpe_2, message=FALSE, warning=FALSE, echo = FALSE}
allRes <- enrichment(tT, organism = "Hs")
print("## BP")
rmarkdown::paged_table(topGO(allRes, ontology = "BP", 
    number = nrow(allRes[allRes[, "Ont"] == "BP", ])))
print("## MF")
rmarkdown::paged_table(topGO(allRes, ontology = "MF", 
    number = nrow(allRes[allRes[, "Ont"] == "MF", ])))
print("## CC")
rmarkdown::paged_table(topGO(allRes, ontology = "CC", 
    number = nrow(allRes[allRes[, "Ont"] == "CC", ])))
```

## FFPE: Association to physico-chemical parameters (contrast 2)

Take the t-values and plot against the GRAVY-Score and Isoelectric point

```{r ffpe_2_phyco, echo = FALSE, eval=FALSE}
t <- tT[, "t"]

## GRAVY score
rD <- rowData(ffpe)
gravy <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateGravyScore(aa))
gravy <- unlist(gravy)
df <- data.frame(gravy = gravy, t = t)
ggplot(df, aes(x = gravy, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(gravy, t, method = "spearman")

## isoelectric point
iep <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateIsoelectricPoint(aa, method = "IPC_protein"))
iep <- unlist(iep)
df <- data.frame(iep = iep, t = t)
ggplot(df, aes(x = iep, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(iep, t, method = "spearman")
```

# Plasma

contrasts:

  - Plasma_AFA - Plasma_TwoPhase_AFA

Upset plot for all condition types.

```{r overlap_plasma_all, echo = FALSE}
MatrixQCvis::upsetCategory(plasma, category = "condition", measured = TRUE)

pdf("upset_protein_plasma_all.pdf")
MatrixQCvis::upsetCategory(plasma, category = "condition", measured = TRUE)
dev.off()
```

## Plasma: Overlap, uniqueness of features


The plot shows the interaction of sets between different variables depending on
their presence. The dots in the UpSet plot specify if the criteria for presence
are fulfilled: Presence is defined by a feature being measured in at least one
sample of a set.

```{r overlap_plasma, echo = FALSE}
plasma <- plasma[, plasma$condition %in% c("Plasma_AFA", "Plasma_TwoPhase_AFA")]
feat_plasma <- extractComb(plasma,
    combination = c("Plasma_AFA", "Plasma_TwoPhase_AFA"),
    measured = TRUE, category = "condition")
MatrixQCvis::upsetCategory(plasma, category = "condition", measured = TRUE)
```

Only continue with the shared features in the following analyses.

```{r plasma_feat, echo = FALSE}
## write also the unique proteins to files
feat <- extractComb(plasma, combination = c("Plasma_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_plasma_Plasma_AFA.txt",
    quote = FALSE, row.names = FALSE)
feat <- extractComb(plasma, combination = c("Plasma_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_plasma_Plasma_TwoPhase.txt",
    quote = FALSE, row.names = FALSE)

## continue only with the overlap
feat <- extractComb(plasma, combination = c("Plasma_AFA", "Plasma_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
plasma <- plasma[feat,]
```


## Plasma: CV

```{r plasma_cv, echo = FALSE}
cv_a <- assay(plasma)[, plasma$condition == "Plasma_AFA"]
cv_b <- assay(plasma)[, plasma$condition == "Plasma_TwoPhase_AFA"]
partial_bundle(MatrixQCvis::cvFeaturePlot(list(Plasma_AFA = cv_a, Plasma_TwoPhase_AFA = cv_b)))

cv_a <- MatrixQCvis::cv(t(cv_a))[[1]]
cv_b <- MatrixQCvis::cv(t(cv_b))[[1]]

## check if one of the two methods has higher means of cv
print("mean: Plasma_AFA")
mean(cv_a, na.rm = TRUE)
print("mean: Plasma_TwoPhase_AFA")
mean(cv_b, na.rm = TRUE)
wilcox.test(cv_a, cv_b, paired = TRUE)
```


## Plasma: Differential expression analysis 

```{r de_plasma, echo = FALSE}
design <- model.matrix(~ 0 + condition, data = colData(plasma))
colnames(design) <- make.names(colnames(design))
fit <- lmFit(object = assay(plasma), design = design, method = "ls")

## create contrasts
contrasts <- makeContrasts(
    AFAvsTwoPhaseAFA = conditionPlasma_AFA - conditionPlasma_TwoPhase_AFA,
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

print("## contrast: plasma_AFA - plasma_TwoPhase_AFA")
tT <- topTable(fit_eB, number = Inf, p.value = p_val, adjust.method = adj,
    coef = "AFAvsTwoPhaseAFA")
tT <- tT[feat, ]
write.table(tT, file = "de_Plasma_plasma_AFA_vs_plasma_TwoPhase_AFA.txt", sep = "\t", quote = FALSE)
rmarkdown::paged_table(tT[order(tT[, "adj.P.Val"]),], options = list(max.print = num))
tT <- cbind(name = rownames(tT), tT)
partial_bundle(volcanoPlot(tT))
```

## Plasma: Enrichment analysis

Run now the GO enrichment tests. Use the raw (unadjusted) values and return
the GO terms for the ontologies Biological Process (BP), Molecular Function (MF) 
and Cellular Component (CC). Perform a Fisher test with the significant features
($\alpha$ < 0.05).
```{r enrichment_plasma, message=FALSE, warning=FALSE, echo = FALSE}
allRes <- enrichment(tT, organism = "Hs")
print("## BP")
rmarkdown::paged_table(topGO(allRes, ontology = "BP", 
    number = nrow(allRes[allRes[, "Ont"] == "BP", ])))
print("## MF")
rmarkdown::paged_table(topGO(allRes, ontology = "MF", 
    number = nrow(allRes[allRes[, "Ont"] == "MF", ])))
print("## CC")
rmarkdown::paged_table(topGO(allRes, ontology = "CC", 
    number = nrow(allRes[allRes[, "Ont"] == "CC", ])))
```

## Plasma: Association to physico-chemical parameters

Take the t-values and plot against the GRAVY-Score and Isoelectric point.

```{r plasma_phyco, echo = FALSE, eval=FALSE}
t <- tT[, "t"]

## GRAVY score
rD <- rowData(plasma)
gravy <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateGravyScore(aa))
gravy <- unlist(gravy)
df <- data.frame(gravy = gravy, t = t)
ggplot(df, aes(x = gravy, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(gravy, t, method = "spearman")

## isoelectric point
iep <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateIsoelectricPoint(aa, method = "IPC_protein"))
iep <- unlist(iep)
df <- data.frame(iep = iep, t = t)
ggplot(df, aes(x = iep, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(iep, t, method = "spearman")
```


# Serum

contrasts:

  - Serum_AFA - Serum_TwoPhase_AFA
  

Upset plot for all condition types.

```{r overlap_serum_all, echo = FALSE}
MatrixQCvis::upsetCategory(serum, category = "condition", measured = TRUE)

pdf("upset_protein_serum_all.pdf")
MatrixQCvis::upsetCategory(serum, category = "condition", measured = TRUE)
dev.off()
```
  
## Serum: Overlap, uniqueness of features

The plot shows the interaction of sets between different variables depending on
their presence. The dots in the UpSet plot specify if the criteria for presence
are fulfilled: Presence is defined by a feature being measured in at least one
sample of a set.

```{r overlap_serum, echo = FALSE}
serum <- serum[, serum$condition %in% c("Serum_AFA", "Serum_TwoPhase_AFA")]
feat_serum <- extractComb(serum, 
    combination = c("Serum_AFA", "Serum_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
MatrixQCvis::upsetCategory(serum, category = "condition", measured = TRUE)
```

Only continue with the shared features in the following analyses.

```{r serum_feat, echo = FALSE}
## write also the unique proteins to files
feat <- extractComb(serum, combination = c("Serum_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_serum_Serum_AFA.txt",
    quote = FALSE, row.names = FALSE)
feat <- extractComb(serum, combination = c("Serum_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
write.table(feat, file = "unique_features_serum_Serum_TwoPhase.txt",
    quote = FALSE, row.names = FALSE)

## continue only with the overlap
feat <- extractComb(serum, combination = c("Serum_AFA", "Serum_TwoPhase_AFA"), 
    measured = TRUE, category = "condition")
serum <- serum[feat,]
```

## Serum: CV

```{r serum_cv, echo = FALSE}
cv_a <- assay(serum)[, serum$condition == "Serum_AFA"]
cv_b <- assay(serum)[, serum$condition == "Serum_TwoPhase_AFA"]
partial_bundle(MatrixQCvis::cvFeaturePlot(list(Serum_AFA = cv_a, Serum_TwoPhase_AFA = cv_b)))

cv_a <- MatrixQCvis::cv(t(cv_a))[[1]]
cv_b <- MatrixQCvis::cv(t(cv_b))[[1]]

## check if one of the two methods has higher means of cv
print("mean: Serum_AFA")
mean(cv_a, na.rm = TRUE)
print("mean: Serum_TwoPhase_AFA")
mean(cv_b, na.rm = TRUE)
wilcox.test(cv_a, cv_b, paired = TRUE)
```

## Serum: Differential expression analysis 

```{r de_serum, echo = FALSE}
design <- model.matrix(~ 0 + condition, data = colData(serum))
colnames(design) <- make.names(colnames(design))
fit <- lmFit(object = assay(serum), design = design, method = "ls")

## create contrasts
contrasts <- makeContrasts(
    AFAvsTwoPhaseAFA = conditionSerum_AFA - conditionSerum_TwoPhase_AFA,
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

print("## contrast: Serum_AFA - Serum_TwoPhase_AFA")
tT <- topTable(fit_eB, number = Inf, p.value = p_val, adjust.method = adj,
    coef = "AFAvsTwoPhaseAFA")
tT <- tT[feat, ]
write.table(tT, file = "de_Serum_Serum_AFA_vs_Serum_TwoPhase_AFA.txt", sep = "\t", quote = FALSE)
rmarkdown::paged_table(tT[order(tT[, "adj.P.Val"]),], options = list(max.print = num))
tT <- cbind(name = rownames(tT), tT)
partial_bundle(volcanoPlot(tT))
```

## Serum: Enrichment analysis

Run now the GO enrichment tests. Use the raw (unadjusted) values and return
the GO terms for the ontologies Biological Process (BP), Molecular Function (MF) and 
Cellular Component (CC). Perform a Fisher test with the significant features
($\alpha$ < 0.05).
```{r enrichment_serum, message=FALSE, warning=FALSE, echo = FALSE}
allRes <- enrichment(tT, organism = "Hs")
print("## BP")
rmarkdown::paged_table(topGO(allRes, ontology = "BP", 
    number = nrow(allRes[allRes[, "Ont"] == "BP", ])))
print("## MF")
rmarkdown::paged_table(topGO(allRes, ontology = "MF", 
    number = nrow(allRes[allRes[, "Ont"] == "MF", ])))
print("## CC")
rmarkdown::paged_table(topGO(allRes, ontology = "CC", 
    number = nrow(allRes[allRes[, "Ont"] == "CC", ])))
```

## Serum: Association to physico-chemical parameters

Take the t-values and plot against the GRAVY-Score and Isoelectric point.

```{r serum_phyco, echo = FALSE, eval=FALSE}
t <- tT[, "t"]

## GRAVY score
rD <- rowData(serum)
gravy <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateGravyScore(aa))
gravy <- unlist(gravy)
df <- data.frame(gravy = gravy, t = t)
ggplot(df, aes(x = gravy, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(gravy, t, method = "spearman")

## isoelectric point
iep <- lapply(rD[rownames(tT), "sequence"], function(aa) calculateIsoelectricPoint(aa, method = "IPC_protein"))
iep <- unlist(iep)
df <- data.frame(iep = iep, t = t)
ggplot(df, aes(x = iep, y = t)) + 
    geom_point(alpha = 0.3) + 
    geom_smooth(method = lm, se = FALSE) +
    ggpubr::stat_cor(method = "spearman") +
    theme_classic()
cor.test(iep, t, method = "spearman")
```


# CV plot for all conditions

```{r cv_all_conditions}
cv_cells_a <- assay(cells)[feat_cells, cells$condition == "Cells_AFA"]
cv_cells_b <- assay(cells)[feat_cells, cells$condition == "Cells_TwoPhase_AFA"]
cv_ff_a <- assay(ff)[feat_ff, ff$condition == "powder_AFA"]
cv_ff_b <- assay(ff)[feat_ff, ff$condition == "powder_TwoPhase_AFA"]
cv_ff_c <- assay(ff)[feat_ff, ff$condition == "Tissue_bulk_AFA"]
cv_ffpe_a <- assay(ffpe)[feat_ffpe, ffpe$condition == "powder_AFA"]
cv_ffpe_b <- assay(ffpe)[feat_ffpe, ffpe$condition == "powder_TwoPhase_AFA"]
cv_ffpe_c <- assay(ffpe)[feat_ffpe, ffpe$condition == "FFPE_AFA"]
cv_plasma_a <- assay(plasma)[feat_plasma, plasma$condition == "Plasma_TwoPhase_AFA"]
cv_plasma_b <- assay(plasma)[feat_plasma, plasma$condition == "Plasma_AFA"]
cv_serum_a <- assay(serum)[feat_serum, serum$condition == "Serum_TwoPhase_AFA"]
cv_serum_b <- assay(serum)[feat_serum, serum$condition == "Serum_AFA"]

## calculate cvs
cv_cells_a <- MatrixQCvis::cv(t(cv_cells_a))[[1]]
cv_cells_b <- MatrixQCvis::cv(t(cv_cells_b))[[1]]
cv_ff_a <- MatrixQCvis::cv(t(cv_ff_a))[[1]]
cv_ff_b <- MatrixQCvis::cv(t(cv_ff_b))[[1]]
cv_ff_c <- MatrixQCvis::cv(t(cv_ff_c))[[1]]
cv_ffpe_a <- MatrixQCvis::cv(t(cv_ffpe_a))[[1]]
cv_ffpe_b <- MatrixQCvis::cv(t(cv_ffpe_b))[[1]]
cv_ffpe_c <- MatrixQCvis::cv(t(cv_ffpe_c))[[1]]
cv_plasma_a <- MatrixQCvis::cv(t(cv_plasma_a))[[1]]
cv_plasma_b <- MatrixQCvis::cv(t(cv_plasma_b))[[1]]
cv_serum_a <- MatrixQCvis::cv(t(cv_serum_a))[[1]]
cv_serum_b <- MatrixQCvis::cv(t(cv_serum_b))[[1]]
 
## create data frames
cv_cells_a <- data.frame(protein = names(cv_cells_a), cv = cv_cells_a, 
    condition = "Cells_AFA", experiment = "cells")
cv_cells_b <- data.frame(protein = names(cv_cells_b), cv = cv_cells_b, 
    condition = "Cells_TwoPhase_AFA", experiment = "cells")
cv_ff_a <- data.frame(protein = names(cv_ff_a), cv = cv_ff_a, 
    condition = "powder_AFA", experiment = "fresh-frozen")
cv_ff_b <- data.frame(protein = names(cv_ff_b), cv = cv_ff_b, 
    condition = "powder_TwoPhase_AFA", experiment = "fresh-frozen")
cv_ff_c <- data.frame(protein = names(cv_ff_c), cv = cv_ff_c, 
    condition = "Tissue_bulk_AFA", experiment = "fresh-frozen")
cv_ffpe_a <- data.frame(protein = names(cv_ffpe_a), cv = cv_ffpe_a, 
    condition = "powder_AFA", experiment = "FFPE")
cv_ffpe_b <- data.frame(protein = names(cv_ffpe_b), cv = cv_ffpe_b, 
    condition = "powder_TwoPhase_AFA", experiment = "FFPE")
cv_ffpe_c <- data.frame(protein = names(cv_ffpe_c), cv = cv_ffpe_c, 
    condition = "FFPE_AFA", experiment = "FFPE")
cv_plasma_a <- data.frame(protein = names(cv_plasma_a), cv = cv_plasma_a, 
    condition = "Plasma_TwoPhase_AFA", experiment = "plasma")
cv_plasma_b <- data.frame(protein = names(cv_plasma_b), cv = cv_plasma_b, 
    condition = "Plasma__AFA", experiment = "plasma")
cv_serum_a <- data.frame(protein = names(cv_serum_a), cv = cv_serum_a, 
    condition = "Serum_TwoPhase_AFA", experiment = "serum")
cv_serum_b <- data.frame(protein = names(cv_serum_b), cv = cv_serum_b, 
    condition = "Serum_AFA", experiment = "serum")

## rbind data frames
cv_df <- rbind(cv_cells_a, cv_cells_b, cv_ff_a, cv_ff_b, cv_ff_c, cv_ffpe_a,
    cv_ffpe_b, cv_ffpe_c, cv_plasma_a, cv_plasma_b, cv_serum_a, cv_serum_b)
cv_df$condition <- factor(cv_df$condition, levels = sort(unique(cv_df$condition)))

g <- ggplot(cv_df) +
    geom_boxplot(aes(x = condition, y = cv)) +
    facet_wrap(~ experiment, scales = "free_x", nrow = 1) +
    theme_classic() + ylab("coefficient of variation") + xlab("") +
    theme(axis.text.x = element_text(angle = 90, 
                                     vjust = 0.5, hjust = 1))
g
ggsave(g, filename = "coefficient_variation_protein_boxplot.pdf")

g <- ggplot(cv_df) +
    geom_violin(aes(x = condition, y = cv)) +
    facet_wrap(~ experiment, scales = "free_x", nrow = 1) +
    theme_classic() + ylab("coefficient of variation") + xlab("") +
    theme(axis.text.x = element_text(angle = 90, 
                                     vjust = 0.5, hjust = 1))
g
ggsave(g, filename = "coefficient_variation_protein_violinplot.pdf")

g <- ggplot(data = cv_df, aes_string(x = "condition", y = "cv", fill = "experiment")) +
    geom_flat_violin(position = position_nudge(x = .2, y = 0), alpha = .8,
        scale = "count") +
    theme_classic() +
    guides(color = "none") +
    scale_fill_manual(values = as.character(wes_palette("Darjeeling1",
        type = "discrete"))) +
    geom_boxplot(width = .1, show.legend = FALSE, outlier.shape = NA,
        alpha = 0.5) +
    xlab("") + ylab("coefficient of variation") +
    facet_wrap(~ experiment, scales = "free_x", nrow = 1) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

g
ggsave(g, filename = "coefficient_variation_protein_violinboxplot.pdf")
```

# Barplot with overlap (%) for different data sets

```{r barplot_overlap, echo = FALSE}
cells_bp <- cells[, cells$condition %in% c("Cells_AFA", "Cells_TwoPhase_AFA")]
ff_bp_1 <- ff[, ff$condition %in% c("powder_AFA", "powder_TwoPhase_AFA")]
ff_bp_2 <- ff[, ff$condition %in% c("Tissue_bulk_AFA", "powder_TwoPhase_AFA")]
ffpe_bp_1 <- ffpe[, ffpe$condition %in% c("powder_AFA", "powder_TwoPhase_AFA")]
ffpe_bp_2 <- ffpe[, ffpe$condition %in% c("powder_TwoPhase_AFA", "FFPE_AFA")]
plasma_bp <- plasma[, plasma$condition %in% c("Plasma_TwoPhase_AFA", "Plasma_AFA")]
serum_bp <- serum[, serum$condition %in% c("Serum_TwoPhase_AFA", "Serum_AFA")]

## extract the features for cells
feat_cells_bp_all <- extractComb(cells_bp, measured = TRUE, category = "condition",
    combination = c("Cells_AFA", "Cells_TwoPhase_AFA"))
feat_cells_bp_other <- extractComb(cells_bp, measured = TRUE, category = "condition",
    combination = c("Cells_AFA"))
feat_cells_bp_tp <- extractComb(cells_bp, measured = TRUE, category = "condition",
    combination = c("Cells_TwoPhase_AFA"))

## extract the features for fresh-frozen (contrast 1)
feat_ff_bp_1_all <- extractComb(ff_bp_1, measured = TRUE, category = "condition",
    combination = c("powder_AFA", "powder_TwoPhase_AFA"))
feat_ff_bp_1_other <- extractComb(ff_bp_1, measured = TRUE, category = "condition",
    combination = c("powder_AFA"))
feat_ff_bp_1_tp <- extractComb(ff_bp_1, measured = TRUE, category = "condition",
    combination = c("powder_TwoPhase_AFA"))

## extract the features for fresh-frozen (contrast 2)
feat_ff_bp_2_all <- extractComb(ff_bp_2, measured = TRUE, category = "condition",
    combination = c("Tissue_bulk_AFA", "powder_TwoPhase_AFA"))
feat_ff_bp_2_other <- extractComb(ff_bp_2, measured = TRUE, category = "condition",
    combination = c("Tissue_bulk_AFA"))
feat_ff_bp_2_tp <- extractComb(ff_bp_2, measured = TRUE, category = "condition",
    combination = c("powder_TwoPhase_AFA"))

## extract the features for FFPE (contrast 1)
feat_ffpe_bp_1_all <- extractComb(ffpe_bp_1, measured = TRUE, category = "condition",
    combination = c("powder_AFA", "powder_TwoPhase_AFA"))
feat_ffpe_bp_1_other <- extractComb(ffpe_bp_1, measured = TRUE, category = "condition",
    combination = c("powder_AFA"))
feat_ffpe_bp_1_tp <- extractComb(ffpe_bp_1, measured = TRUE, category = "condition",
    combination = c( "powder_TwoPhase_AFA"))

## extract the features for FFPE (contrast 2)
feat_ffpe_bp_2_all <- extractComb(ffpe_bp_2, measured = TRUE, category = "condition",
    combination = c("powder_TwoPhase_AFA", "FFPE_AFA"))
feat_ffpe_bp_2_other <- extractComb(ffpe_bp_2, measured = TRUE, category = "condition",
    combination = c("FFPE_AFA"))
feat_ffpe_bp_2_tp <- extractComb(ffpe_bp_2, measured = TRUE, category = "condition",
    combination = c("powder_TwoPhase_AFA"))

## extract the features for plasma 
feat_plasma_bp_all <- extractComb(plasma_bp, measured = TRUE, category = "condition",
    combination = c("Plasma_TwoPhase_AFA", "Plasma_AFA"))
feat_plasma_bp_other <- extractComb(plasma_bp, measured = TRUE, category = "condition",
    combination = c("Plasma_AFA"))
feat_plasma_bp_tp <- extractComb(plasma_bp, measured = TRUE, category = "condition",
    combination = c("Plasma_TwoPhase_AFA"))

## extract the features for serum
feat_serum_bp_all <- extractComb(serum_bp, measured = TRUE, category = "condition",
    combination = c("Serum_TwoPhase_AFA", "Serum_AFA"))
feat_serum_bp_other <- extractComb(serum_bp, measured = TRUE, category = "condition",
    combination = c("Serum_AFA"))
feat_serum_bp_tp <- extractComb(serum_bp, measured = TRUE, category = "condition",
    combination = c("Serum_TwoPhase_AFA"))

## calculate the percentage of shared and unique features and write to data.frame
n_cells <- 100 / length(c(feat_cells_bp_all, feat_cells_bp_other, feat_cells_bp_tp))
n_ff_1 <- 100 / length(c(feat_ff_bp_1_all, feat_ff_bp_1_other, feat_ff_bp_1_tp))
n_ff_2 <- 100 / length(c(feat_ff_bp_2_all, feat_ff_bp_2_other, feat_ff_bp_2_tp))
n_ffpe_1 <- 100 / length(c(feat_ffpe_bp_1_all, feat_ffpe_bp_1_other, feat_ffpe_bp_1_tp))
n_ffpe_2 <- 100 / length(c(feat_ffpe_bp_2_all, feat_ffpe_bp_2_other, feat_ffpe_bp_2_tp))
n_plasma <- 100 / length(c(feat_plasma_bp_all, feat_plasma_bp_other, feat_plasma_bp_tp))
n_serum <- 100 / length(c(feat_serum_bp_all, feat_serum_bp_other, feat_serum_bp_tp))
df_bp <- data.frame(
    rbind(c(length(feat_cells_bp_all) * n_cells, "cells", "shared"),
        c(length(feat_cells_bp_other) * n_cells, "cells", "other"),
        c(length(feat_cells_bp_tp) * n_cells, "cells", "two phase"),
        c(length(feat_ff_bp_1_all) * n_ff_1, "fresh-frozen (powder)", "shared"),
        c(length(feat_ff_bp_1_other) * n_ff_1, "fresh-frozen (powder)", "other"),
        c(length(feat_ff_bp_1_tp) * n_ff_1, "fresh-frozen (powder)", "two phase"),  
        c(length(feat_ff_bp_2_all) * n_ff_2, "fresh-frozen (bulk)", "shared"),
        c(length(feat_ff_bp_2_other) * n_ff_2, "fresh-frozen (bulk)", "other"),
        c(length(feat_ff_bp_2_tp) * n_ff_2, "fresh-frozen (bulk)", "two phase"),
        c(length(feat_ffpe_bp_1_all) * n_ffpe_1, "FFPE (powder)", "shared"),
        c(length(feat_ffpe_bp_1_other) * n_ffpe_1, "FFPE (powder)", "other"),
        c(length(feat_ffpe_bp_1_tp) * n_ffpe_1, "FFPE (powder)", "two phase"),  
        c(length(feat_ffpe_bp_2_all) * n_ffpe_2, "FFPE (bulk)", "shared"),
        c(length(feat_ffpe_bp_2_other) * n_ffpe_2, "FFPE (bulk)", "other"),
        c(length(feat_ffpe_bp_2_tp) * n_ffpe_2, "FFPE (bulk)", "two phase"),
        c(length(feat_plasma_bp_all) * n_plasma, "plasma", "shared"),
        c(length(feat_plasma_bp_other) * n_plasma, "plasma", "other"),
        c(length(feat_plasma_bp_tp) * n_plasma, "plasma", "two phase"),
        c(length(feat_serum_bp_all) * n_serum, "serum", "shared"),
        c(length(feat_serum_bp_other) * n_serum, "serum", "other"),
        c(length(feat_serum_bp_tp) * n_serum, "serum", "two phase")))
colnames(df_bp) <- c("percent", "material", "type")
df_bp$percent <- as.numeric(df_bp$percent)
df_bp$type <- factor(df_bp$type, levels = c("other", "two phase", "shared"))

g <- ggplot(df_bp, aes(y = percent, x = material, col = type, fill = type)) + 
    geom_bar(stat = "identity") + 
    scale_fill_manual(values = as.character(wes_palette("FantasticFox1", 
        type = "discrete"))) +
    scale_color_manual(values = as.character(wes_palette("FantasticFox1",
        type = "discrete"))) +
    ##facet_wrap(~ material, nrow = 1, scales = "free_x") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 90))
g
ggsave(g, filename = "barplot_overlap_protein.pdf")
```
