---
title: "Comparison of Extraction Methods - differential expression of metabolites"
subtitle: "Thorax"
author:
    - Thomas Naake^[European Molecular Biology Laboratory, Meyerhofstrasse 1, 69117 Heidelberg, Germany]
fig_width: 15
fig_height: 10
fontsize: 12pt
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: TRUE
    theme: united
    number_sections: true
    highlight: tango
---

```{r env, include=FALSE, echo=FALSE, cache=FALSE}
library("knitr")
library("ggplot2")
library("tidyverse")
knitr::opts_chunk$set(stop_on_error = 1L, fig.pos = "ht", dev = "png")
suppressPackageStartupMessages(library("wesanderson"))
suppressPackageStartupMessages(library("MatrixQCvis"))
suppressPackageStartupMessages(library("PhysicoChemicalPropertiesProtein"))
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("limma"))
knitr::opts_knit$set(root.dir = "~/GitLab/20211202_extraction_method/")
```

# Load data 

```{r load_data}
se <- biocrates("2022-02-22_Conc_raw_smartcare 10 patient thorax cohort_hmdb.xlsx")
se$tissue <- ""
se$tissue[grep(x = se$Sample.Description, pattern = "_TU")] <- "TU"
se$tissue[grep(x = se$Sample.Description, pattern = "_NG")] <- "NG"
se$individual <- stringr::str_remove(se$Sample.Description, pattern = "_TU|_NG")
```

# Load data and filter features

```{r metIDQ_get_high_quality_features_fct, eval = TRUE, echo = FALSE}
## Define a helper function to remove low-quality features. The function 
## `cellColor` is used within the function `metIDQ_get_high_quality_features`.
setwd("C:/Users/naake/Documents/GitLab/20211202_extraction_method/")

#' @name cellColor
#' 
#' @title cellColor - helper function for metIDQ_get_high_quality_features
cellColor <- function(style) {
        fg  <- style$getFillForegroundXSSFColor()
        rgb <- tryCatch(fg$getRgb(), error = function(e) "")
        rgb <- paste(rgb, collapse = "")
        return(rgb)
}

#' @name metIDQ_get_high_quality_features
#'
#' @title Get a vector with the high-quality features to retain
metIDQ_get_high_quality_features <- function(file, threshold = 0.66) {
    ## get the background color
    wb <- xlsx::loadWorkbook(file = file)
    sheet1 <- xlsx::getSheets(wb)[[1]]
    rows <- xlsx::getRows(sheet1)
    cells <- xlsx::getCells(rows)
    styles <- sapply(cells, xlsx::getCellStyle)
    bg <- sapply(styles, cellColor) 
    
    ## use values to get the indices where the actual data is stored
    values <- sapply(cells, xlsx::getCellValue)

    ## convert bg and values to matrix that it has the same dimension as before
    if (all(names(bg) != names(values))) stop("dimensions do not match")
    row_index <- as.numeric(unlist(lapply(strsplit(names(bg), split = "[.]"), "[", 1)))
    col_index <- as.numeric(unlist(lapply(strsplit(names(bg), split = "[.]"), "[", 2)))
    mat_values <- mat_bg <- matrix("", ncol = max(col_index), nrow = max(row_index))
    for (i in 1:length(bg)) {
        mat_bg[row_index[i], col_index[i]] <- bg[i]
        mat_values[row_index[i], col_index[i]] <- values[i]
    }
    
    ## obtain the indices of the cells where the values/background colors are 
    ## stored in
    row_inds <- which(!is.na(mat_values[, 1]))
    row_inds <- row_inds[!row_inds %in% 1:27]
    col_inds <- rbind(which(mat_values == "C0", arr.ind = TRUE),
                    which(mat_values == "Choline", arr.ind = TRUE))
    colnames(mat_values) <- colnames(mat_bg) <- mat_values[col_inds[1, "row"], ]
    col_inds <- seq(col_inds[1, "col"], col_inds[2, "col"])
    
    ## iterate through the columns of mat_bg and check the color values 
    ## against the QC of MetIDS## "00cd66" == green, "87ceeb" == lightblue
    valid <- apply(mat_bg[row_inds, col_inds], 2, 
        function(x) sum(x %in% c("00cd66", "87ceeb"))) 

    ## require that at least threshold*100% values per metabolite are 
    ## "green"/"lightblue", 
    valid / length(row_inds)  > threshold
}
```

Run the function `metIDQ_get_high_quality_features` and only retain the 
features that have more than 2/3 of "green"/"lightblue" values.

```{r filter}
file <- "2022-02-22_Conc_raw_smartcare 10 patient thorax cohort_hmdb.xlsx"
features <- metIDQ_get_high_quality_features(file, threshold = 2/3)
se <- se[make.names(names(which(features))), ]
```


# QC

```{r shinyQC, eval = FALSE}
shinyQC(se)
```

Exclude the following samples:
`PBS` samples and `QC` samples.


```{r exclude_samples}
exclude <- c("PBS_1", "PBS_2", "PBS_3", "Quant 500_QC1_1", "Quant 500_QC2_2",
    "Quant 500_QC2_3", "Quant 500_QC2_4", "Quant 500_QC3_5")
se <- MatrixQCvis:::selectSampleSE(se, 
    selection = exclude,
    mode = "exclude")
```


# Data Transformation

Perform `log`-transformation on the data set.

```{r data_transformation}
a <- assay(se)
a[a <= 0] <- NA
a_t <- transformAssay(a, method = "log")
se <- MatrixQCvis:::updateSE(se = se, a = a_t)

saveRDS(se, file = "SummarizedExperiment_extraction_method_cohort_metabolomics.RDS")
```

# Run the linear model

```{r}
cD <- colData(se)
cD <- colData(se)
design <- model.matrix(~ 0 + tissue, data = cD)
colnames(design) <- make.names(colnames(design))
cor <- duplicateCorrelation(assay(se), design, block=cD$individual)
fit <- lmFit(object = assay(se), design = design, block = cD$individual,
    correlation = cor$consensus)

## create contrasts
contrasts <- makeContrasts(
    TUvsNG = ( tissueTU - tissueNG),
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

## set parameters for differential expression
num <- Inf
p_val <- 1
adj <- "BH"
```  


## TU vs. NG

We test here the DE proteins between NG and TU. 

```{r autoSP3vsMTBE_SP3}
tT <- topTable(fit_eB, number = num, p.value = p_val, adjust.method = adj,
    coef = "TUvsNG")
rmarkdown::paged_table(tT)
sum(tT$adj.P.Val < 0.05, na.rm = TRUE)
tT <- cbind(name = rownames(tT), tT)
volcanoPlot(tT)

write.table(tT, file = "metabolomics_DE_t_TUvsNG.txt", 
    quote = FALSE, sep = "\t")
```

## Individuals as covariates

```{r}
cD <- colData(se)
design <- model.matrix(~ 0 + tissue + individual, data = cD)
fit <- lmFit(object = assay(se), design = design)

## create contrasts
contrasts <- makeContrasts(
    TUvsNG = (tissueTU - tissueNG),
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

## set parameters for differential expression
num <- Inf
p_val <- 1
adj <- "BH"

## get the features for autoSP3
tT_covariate <- topTable(fit_eB, number = num, p.value = p_val, 
    adjust.method = adj, coef = "TUvsNG")

cor(x = tT_covariate$t[order(rownames(tT_covariate))], y = tT$t[order(rownames(tT))], 
    use = "pairwise.complete.obs")
```

