---
title: "Comparison of Extraction Methods (proteins)"
subtitle: "Thorax"
author:
    - Thomas Naake^[European Molecular Biology Laboratory, Meyerhofstrasse 1, 69117 Heidelberg, Germany]
fig_width: 15
fig_height: 10
fontsize: 12pt
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: TRUE
    theme: united
    number_sections: true
    highlight: tango
---

```{r env, include=FALSE, echo=FALSE, cache=FALSE}
library("knitr")
library("ggplot2")
knitr::opts_chunk$set(stop_on_error = 1L, fig.pos = "ht", dev = "png")
suppressPackageStartupMessages(library("wesanderson"))
suppressPackageStartupMessages(library("MatrixQCvis"))
suppressPackageStartupMessages(library("MatrixQCvisUtils"))
suppressPackageStartupMessages(library("PhysicoChemicalPropertiesProtein"))
suppressPackageStartupMessages(library("tidyverse"))
suppressPackageStartupMessages(library("limma"))
library("ggbeeswarm")
knitr::opts_knit$set(root.dir = "~/GitLab/20211202_extraction_method/")
```

# Load data 

```{r load_data}
se <- maxQuant("proteinGroups_thorax.xlsx", intensity = "LFQ", type = "xlsx", 
    sheet = "filtered")
## load annotation
cD <- openxlsx::read.xlsx("proteinGroups_thorax.xlsx", sheet = "annotation_filtered")
cD[, "Sample_IDs"] <- make.names(cD[, "Sample_IDs"])
colnames(cD)[colnames(cD) == "Sample_IDs"] <- "name"

## truncate se and write cD to colData(se)
se <- se[, cD$name]
se@colData <- cD %>% DataFrame()
colnames(se) <- se$name

## truncate colnames
colnames(se) <- stringr::str_replace(colnames(se), 
    "LFQ.intensity.1_KL_ProtMet_IO_40min_DDA_", "ProtMet_")
colnames(se) <- stringr::str_replace(colnames(se), 
    "LFQ.intensity.5_KL_nWorkFlow_IO_40min_DDA_", "nWorkFlow_")
colnames(se) <- stringr::str_replace(colnames(se), 
    "LFQ.intensity.5_KL_nWorkFlowt_IO_40min_DDA_", "nWorkFlow")

se$name <- colnames(se)
```

# QC

```{r shinyQC, eval = FALSE}
shinyQC(se)
```

Exclude the following samples:
`ProtMet_8U04FR_NG_14_B10_1_6020`, `ProtMet_IJ17TV_NG_12_B8_1_6016`.

```{r exclude_samples}
se <- MatrixQCvis:::selectSampleSE(se, 
    selection = c("ProtMet_8U04FR_NG_14_B10_1_6020", "ProtMet_IJ17TV_NG_12_B8_1_6016"),
    mode = "exclude")
```

## Translate the protein IDs

Translate Uniprot into Symbol.

```{r, message=FALSE, warning=FALSE}
library("org.Hs.eg.db")
uniprots <- Rkeys(org.Hs.egUNIPROT)
dict <- AnnotationDbi::select(org.Hs.eg.db, uniprots, "SYMBOL", "UNIPROT")
uniprot <- dict$UNIPROT
symbol <- dict$SYMBOL
names(symbol) <- uniprot

## rename
rownames(se) <- unlist(lapply(rownames(se), 
    function(x) paste(symbol[strsplit(x, split = ";")[[1]]], collapse = ";")))

## remove the features that have no corresponding Symbol
exclude_pattern <- c("NA", "NA;NA", "NA;NA;NA", "NA;NA;NA;NA", 
    "NA;NA;NA;NA;NA", "NA;NA;NA;NA;NA;NA", "NA;NA;NA;NA;NA;NA;NA")
se <- se[!rownames(se) %in% exclude_pattern, ]
```

# Data Filtering, Transformation, and Imputation

Perform `log`-transformation on the data set.

```{r data_transformation}
## keep features with more than ca. 60% (18/35) measured values
se <- se[rowSums(!is.na(assay(se))) >= 18 , ]
dim(se)

## transformation and imputation
a <- assay(se)
a_t <- transformAssay(a, method = "log")
se <- MatrixQCvis:::updateSE(se = se, assay = a_t)

saveRDS(se, file = "SummarizedExperiment_extraction_method_cohort_proteomics.RDS")
```

# Run the linear model

```{r}
cD <- colData(se)
cD$Type_Processing <- paste(cD$Type, cD$Processing, sep = "_")
cD$Individual <- stringr::str_remove(cD$Pseudonym, pattern = "_NG|_TU")
design <- model.matrix(~ 0 + Type_Processing, data = cD)
colnames(design) <- make.names(colnames(design))
cor <- limma::duplicateCorrelation(assay(se), design, block=cD$Individual)
fit <- lmFit(object = assay(se), design = design, block = cD$Individual,
    correlation = cor$consensus)

## create contrasts
contrasts <- makeContrasts(
    autoSP3vsMTBE_SP3 = (Type_ProcessingTU_autoSP3 - Type_ProcessingNG_autoSP3)/2 -
        (Type_ProcessingTU_MTBE_SP3 - Type_ProcessingNG_MTBE_SP3)/2,
    TUvsNG = (Type_ProcessingTU_autoSP3 - Type_ProcessingTU_MTBE_SP3)/2 -
        (Type_ProcessingNG_autoSP3 - Type_ProcessingNG_MTBE_SP3)/2, ## identical to autoSP3vsMTBE_SP3 
    NG_autoSP3vsMTBE_SP3 = Type_ProcessingNG_autoSP3 - Type_ProcessingNG_MTBE_SP3,
    TU_autoSP3vsMTBE_SP3 = Type_ProcessingTU_autoSP3 - Type_ProcessingTU_MTBE_SP3,
    autoSP3 = Type_ProcessingTU_autoSP3 - Type_ProcessingNG_autoSP3,
    MTBE_SP3 = Type_ProcessingTU_MTBE_SP3 - Type_ProcessingNG_MTBE_SP3,
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

## set parameters for differential expression
num <- Inf
p_val <- 1
adj <- "BH"
```  


## autoSP3 vs. MTBE_SP3

We test here the DE proteins between NG and TU and compare the results that
we get from the two extraction methods. 

Ideally, this should result in few DE proteins. This would indicate that the 
methods yield similar results.

```{r autoSP3vsMTBE_SP3}
tT <- topTable(fit_eB, number = num, p.value = p_val, adjust.method = adj,
    coef = "autoSP3vsMTBE_SP3")
rmarkdown::paged_table(tT)
sum(tT[, "adj.P.Val"] < 0.05, na.rm = TRUE)
tT <- cbind(name = rownames(tT), tT)
volcanoPlot(tT)
tT_autoSP3vsMTBE <- tT

write.table(tT, file = "proteomics_DE_t_autoSP3vsMTBE_SP3.txt", 
    quote = FALSE, sep = "\t")
```

## autoSP3 vs. MTBE_SP3 (NG)

We test here the DE proteins from the two extraction methods looking only at 
NG. 

Ideally, this should result in few DE proteins. This would indicate that the 
methods yield similar results.

```{r NG_autoSP3vsMTBE_SP3}
tT <- topTable(fit_eB, number = num, p.value = p_val, adjust.method = adj,
    coef = "NG_autoSP3vsMTBE_SP3")
rmarkdown::paged_table(tT)
sum(tT[, "adj.P.Val"] < 0.05, na.rm = TRUE)
tT <- cbind(name = rownames(tT), tT)
volcanoPlot(tT)
tT_NG <- tT

write.table(tT, file = "proteomics_DE_t_NG_autoSP3vsMTBE_SP3.txt", 
    quote = FALSE, sep = "\t")
```


## autoSP3 vs. MTBE_SP3 (TU)

We test here the DE proteins from the two extraction methods looking only at 
TU 

Ideally, this should result in few DE proteins. This would indicate that the 
methods yield similar results.

```{r TU_autoSP3vsMTBE_SP3}
tT <- topTable(fit_eB, number = num, p.value = p_val, adjust.method = adj,
    coef = "TU_autoSP3vsMTBE_SP3")
rmarkdown::paged_table(tT)
sum(tT[, "adj.P.Val"] < 0.05, na.rm = TRUE)
tT <- cbind(name = rownames(tT), tT)
volcanoPlot(tT)
tT_TU <- tT

write.table(tT, file = "proteomics_DE_t_TU_autoSP3vsMTBE_SP3.txt", 
    quote = FALSE, sep = "\t")
```


Ok, this is interesting, when looking at the individual tissues, there are quite
many differences, but the differences "vanish" when we combine the different
conditions (autoSP3 vs. MTBE_SP3).

## Overlap between DE proteins

We will continue and check the overlap of DE proteins between the contrasts

    - autoSP3 vs. MTBE_SP3 (NG)
    - autoSP3 vs. MTBE_SP3 (TU)
    - autoSP3 vs. MTBE_SP3 (TU-NG)
    
```{r de_overlap_1}
## only continue with the significant ones
NG_sign <- tT_NG[tT_NG$adj.P.Val < 0.05 & !is.na(tT_NG$adj.P.Val), ]
dim(NG_sign)
TU_sign <- tT_TU[tT_TU$adj.P.Val < 0.05 & !is.na(tT_TU$adj.P.Val), ]
dim(TU_sign)
autoSP3vsMTBE_sign <- tT_autoSP3vsMTBE[tT_autoSP3vsMTBE$adj.P.Val < 0.05 & !is.na(tT_autoSP3vsMTBE$adj.P.Val), ]
dim(autoSP3vsMTBE_sign)

## create list
l <- list(NG = rownames(NG_sign), TU = rownames(TU_sign), 
    autoSP3vsMTBE_SP3 = rownames(autoSP3vsMTBE_sign))
UpSetR::upset(UpSetR::fromList(l), order.by = "freq")

pdf("UpSet_NAT_TT_autoSP3vsMTBESP3.pdf")
UpSetR::upset(UpSetR::fromList(l), order.by = "freq")
dev.off()
```


We will continue and check the overlap of DE proteins between the contrasts:
    
    - TU vs. NG (autoSP3)
    - TU vs. NG (MTBE_SP3)
    
```{r de_overlap}
## autoSP3
tT_autoSP3 <- topTable(fit_eB, number = num, p.value = p_val, adjust.method = adj,
    coef = "autoSP3")
sum(tT_autoSP3$adj.P.Val < 0.05, na.rm = TRUE)
tT_autoSP3 <- cbind(name = rownames(tT_autoSP3), tT_autoSP3)
volcanoPlot(tT_autoSP3)
write.table(tT_autoSP3, file = "proteomics_DE_t_TUvsNG_autoSP3.txt", 
    quote = FALSE, sep = "\t")

## MTBE_SP3
tT_MTBE_SP3 <- topTable(fit_eB, number = num, p.value = p_val, adjust.method = adj,
    coef = "MTBE_SP3")
rmarkdown::paged_table(tT_MTBE_SP3)
sum(tT_MTBE_SP3$adj.P.Val < 0.05, na.rm = TRUE)
tT_MTBE_SP3 <- cbind(name = rownames(tT_MTBE_SP3), tT_MTBE_SP3)
volcanoPlot(tT_MTBE_SP3)

write.table(tT_MTBE_SP3, file = "proteomics_DE_t_TUvsNG_MTBE_SP3.txt", 
    quote = FALSE, sep = "\t")

## only continue with the significant ones 
autoSP3_sign <- tT_autoSP3[tT_autoSP3$adj.P.Val < 0.05 & !is.na(tT_autoSP3$adj.P.Val), ]
MTBE_SP3_sign <- tT_MTBE_SP3[tT_MTBE_SP3$adj.P.Val < 0.05  & !is.na(tT_MTBE_SP3$adj.P.Val), ]

## create list
l <- list(autoSP3 = autoSP3_sign$ID, MTBE_SP3 = MTBE_SP3_sign$ID)
UpSetR::upset(UpSetR::fromList(l), order.by = "freq")

pdf("UpSet_NATvsTT_autoSP3_MTBESP3.pdf")
UpSetR::upset(UpSetR::fromList(l), order.by = "freq")
dev.off()

## Can we say anything about the magnitude of the difference? E.g. the
## overlapping 1004 proteins show on average a larger fold change than those
## detected by only one of the methods? Show in a box plot.
mtbe_auto_shared <- l$autoSP3[l$autoSP3 %in% l$MTBE_SP3]
mtbe_unique <- l$MTBE_SP3[!(l$MTBE_SP3 %in% l$autoSP3)]
auto_unique <- l$autoSP3[!(l$autoSP3 %in% l$MTBE_SP3)]

df_mtbe_shared <- data.frame(
    logFC = tT_MTBE_SP3[tT_MTBE_SP3$ID %in% mtbe_auto_shared, "logFC"],
    set = "shared", type = "MTBE-SP3")
df_auto_shared <- data.frame(
    logFC = tT_autoSP3[tT_autoSP3$ID %in% mtbe_auto_shared, "logFC"],
    set = "shared", type = "autoSP3")
df_mtbe_unique <- data.frame(
    logFC = tT_MTBE_SP3[tT_MTBE_SP3$ID %in% mtbe_unique, "logFC"],
    set = "unique", type = "MTBE-SP3")
df_auto_unique <- data.frame(
    logFC = tT_autoSP3[tT_autoSP3$ID %in% auto_unique, "logFC"],
    set = "unique", type = "autoSP3")
df_all <- rbind(df_mtbe_shared, df_auto_shared, df_mtbe_unique, df_auto_unique)

g <- ggplot(df_all) +
    geom_beeswarm(aes(y = logFC, x = type), size = 0.5, alpha = 0.5) +
    facet_wrap(~ set + type, ncol = 4, scales = "free_x") +
    theme_classic()
ggsave(filename = "BeeSwarm_NATvsTT_autoSP3_MTBESP3.pdf", plot = g)

df_tmp <- df_all
df_tmp$logFC <- abs(df_tmp$logFC)
wilcox.test(
    df_tmp[df_tmp$set == "shared" & df_tmp$type == "autoSP3", "logFC"],
    df_tmp[df_tmp$set == "unique" & df_tmp$type == "autoSP3", "logFC"],
    alternative = "greater")
wilcox.test(
    df_tmp[df_tmp$set == "shared" & df_tmp$type == "MTBE-SP3", "logFC"],
    df_tmp[df_tmp$set == "unique" & df_tmp$type == "MTBE-SP3", "logFC"],
    alternative = "greater")

## TU vs NG. (taking into account baseline autoSP3/MTBE-SP3)
tT <- topTable(fit_eB, number = num, p.value = p_val, adjust.method = adj,
    coef = "TUvsNG")
rmarkdown::paged_table(tT)
tT <- cbind(name = rownames(tT), tT)
volcanoPlot(tT)
tT_TUvsNG <- tT

## plot t-values
df <- data.frame(vals_x = tT_autoSP3[rownames(tT_autoSP3), "t"], 
        vals_y = tT_MTBE_SP3[rownames(tT_autoSP3), "t"],
        vals_z = tT_TUvsNG[rownames(tT_autoSP3), "t"])
rownames(df) <- rownames(tT_autoSP3)

## t-values(autoSP3) vs t-values (MTBE-SP3)
g <- ggplot(df, aes_string(x = "vals_x", y = "vals_y")) +
    geom_point(aes(alpha = 0.7)) +
    xlab("t-values (autoSP3)") + ylab("t-values (MTBE-SP3)") +
    ggpubr::stat_cor(method = "spearman", cor.coef.name = "rho") +
    ggtitle("NAT vs. TT") +
    theme_classic() +
    theme(legend.position = "none")
g 
ggsave(g, filename = "scatter_NATvsTT_autoSP3_MTBESP3.pdf")
cor.test(tT_autoSP3[rownames(tT_autoSP3), "t"],
    tT_MTBE_SP3[rownames(tT_autoSP3), "t"], method = "spearman")

## t-values(autoSP3) vs t-values (AutoSP3vsMTBE-SP3)
g <- ggplot(df, aes_string(x = "vals_x", y = "vals_z")) +
    geom_point(aes(alpha = 0.7)) +
    xlab("t-values (autoSP3)") + ylab("t-values (autoSP3/MTBE-SP3)") +
    ggpubr::stat_cor(method = "spearman", cor.coef.name = "rho") +
    ggtitle("NAT vs. TT") +
    theme_classic() +
    theme(legend.position = "none")
g 
ggsave(g, filename = "scatter_NATvsTT_autoSP3_autoSP3vsMTBESP3.pdf")

## t-values(MTBE-SP3) vs t-values (AutoSP3vsMTBE-SP3)
g <- ggplot(df, aes_string(x = "vals_y", y = "vals_z")) +
    geom_point(aes(alpha = 0.7)) +
    xlab("t-values (MTBE-SP3)") + ylab("t-values (autoSP3/MTBE-SP3)") +
    ggpubr::stat_cor(method = "spearman", cor.coef.name = "rho") +
    ggtitle("NAT vs. TT") +
    theme_classic() +
    theme(legend.position = "none")
g 
ggsave(g, filename = "scatter_NATvsTT_MTBESP3_autoSP3vsMTBESP3.pdf")


```

## Individuals as covariates

```{r}
cD <- colData(se)
cD$Type_Processing <- paste(cD$Type, cD$Processing, sep = "_")
cD$Individual <- stringr::str_remove(cD$Pseudonym, pattern = "_NG|_TU")
design <- model.matrix(~ 0 + Type_Processing + Individual, data = cD)
fit <- lmFit(object = assay(se), design = design)

## create contrasts
contrasts <- makeContrasts(
    autoSP3 = Type_ProcessingTU_autoSP3 - Type_ProcessingNG_autoSP3,
    MTBE_SP3 = Type_ProcessingTU_MTBE_SP3 - Type_ProcessingNG_MTBE_SP3,
    levels = design)
fit_c <- contrasts.fit(fit, contrasts)
fit_eB <- eBayes(fit_c)

## set parameters for differential expression
num <- Inf
p_val <- 1
adj <- "BH"

## get the features for autoSP3
tT_autoSP3_covariate <- topTable(fit_eB, number = num, p.value = p_val, 
    adjust.method = adj, coef = "autoSP3")

cor(x = tT_autoSP3_covariate$t[order(rownames(tT_autoSP3_covariate))],
    y = tT_autoSP3$t[order(rownames(tT_autoSP3))], 
    use = "pairwise.complete.obs")

## get the features for MTBE_SP3
tT_MTBE_SP3_covariate <- topTable(fit_eB, number = num, p.value = p_val, 
    adjust.method = adj, coef = "MTBE_SP3")

cor(x = tT_MTBE_SP3_covariate$t[order(rownames(tT_MTBE_SP3_covariate))],
    y = tT_MTBE_SP3$t[order(rownames(tT_MTBE_SP3))], 
    use = "pairwise.complete.obs")
```


